---
title: "Examples"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
devtools::load_all("..")
```

```{r options, echo=FALSE, message=FALSE}
simrOptions(progress=FALSE)
```

First we'll set the default number of simulations to a lower value. This will allow us to run the examples more quickly.

```{r}
simrOptions(nsim=25)
```

## Binomial GLMM with categorical predictor

The first example comes from the online help for `glmer`. The data frame `cbpp` contains data on contagious bovine pleuropneumonia. Note that the response is specified using `cbind` --- `simr` expects a binomial model to be in this form.

```{r}
gm1 <- glmer(cbind(incidence, size - incidence) ~ period + (1 | herd), data = cbpp, family = binomial)
summary(gm1)
```

Note that `period` is a categorical variable with 4 levels. `simr` will default to a likelihood ratio test comparing `gm1` to the reduced model without the `period` term.

```{r}
powerSim(gm1)
```

If you were (for some reason) especially interested in the power for the dummy variable `period2` you could use a Z-test. This tests the value `Pr(>|z|)` reported in the summary above.

```{r}
powerSim(gm1, fixed("period2", "z"))
```

Suppose your model also has a continuous predictor. You can use `fixed` to choose which fixed effect to apply tests to. 

```{r}
gm2 <- glmer(cbind(incidence, size - incidence) ~ period + size + (1 | herd), data = cbpp, family = binomial)
fixef(gm2)['size'] <- 0.05
powerSim(gm2, fixed("size"))
```

## Models with interaction terms

### Cake

This example uses the `cake` dataset.

```{r}
fm1 <- lmer(angle ~ recipe * temp + (1|recipe:replicate), cake, REML= FALSE)
summary(fm1)
```

Main effects should not be tested when they appear in an interaction term. Using the `compare` function, we can specify a comparison with a simpler model.

```{r}
powerSim(fm1, compare(angle ~ recipe + temp + (1|recipe:replicate)))
```

### Budworms

We can do the same thing with the `budworm` data in the `pbkrtest` package.

```{r}
data(budworm)
bw1 <- glm(cbind(ndead, ntotal-ndead) ~ dose*sex, family="binomial", data=budworm)
```

Of course we don't want to retype the `cbind` boilerplate:

```{r}
powerSim(bw1, compare(. ~ dose + sex))
```

Since `dose` is continous and `sex` is binary we can also use a Z-test on the single interaction term.

```{r}
powerSim(bw1, fixed("dose:sexmale", "z"))
```

## Single random effects

The `random` function gives you access to tests from the `RLRsim` package. No additional arguments are needed for a single random effect.

```{r}
re1 <- lmer(Yield ~ 1|Batch, Dyestuff)
powerSim(re1, random())
```

### Note about errors during simulation

Some tests will occasionally fail on simulated data. When this happens `simr` calculates power based on the iterations that completed without generating errors.

```{r}
powerSim(re1, random(), seed=17)
```

Rather than stopping part-way through am analysis, `simr` traps and records errors. You can access these records using `$warnings` and `$errors`.

```{r}
lastResult()$warnings
lastResult()$errors
```

## Multiple random effects

Where the model has multiple random effects, `compare` can be used to test alternative specifications. 

```{r}
fm1 <- lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
```

```{r}
powerSim(fm1, compare(Reaction ~ Days + (1 | Subject)))
```

